---
title: "Adult Dataset"
author: "KidneyA11"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# load library
#aBiocManager::install("preprocessCore")
library(GEOquery)
library(dplyr)
library(tidyverse)
library(limma)
library(reshape2)
library(DT)
library(stringr)
library(preprocessCore)
library(sva)
library(cvTools)      # Cross-validation
library(e1071)        # Support vector machines
library(randomForest) # Random forests
library(viridis)      # Viridis colour scheme
library(plotly)       # Interactive plots
```

# Load Datasets
```{r warning=FALSE}
# load datasets for Adult
gseAdultA <- getGEO("GSE46474")
gseAdultA_raw <- gseAdultA[[1]] # Extract the expression data

gseAdultB <- getGEO("GSE1563")
gseAdultB_raw <- gseAdultB[[1]] # Extract the expression data

gseAdultC <- getGEO("GSE14328")
gseAdultC_raw <- gseAdultC[[1]] # Extract the expression data
```

## Adult: A. GSE46474

### Basic statistics

```{r}
#pData(gseAdultA_raw)
#colnames(pData(gseAdultA_raw))
gseAdultA_raw$Outcome <- ifelse(grepl("AR", pData(gseAdultA_raw)$characteristics_ch1.5), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultA_raw$Outcome)
dim(gseAdultA_raw)
```

### Cleaning

```{r}
# preprocessing
fDataAdultA_raw <- fData(gseAdultA_raw) # Extract the gene information
#colnames(fDataAdultA_raw)
#head(fDataAdultA_raw[,"Gene Symbol"])

design_AdultA <- model.matrix(~Outcome, data = pData(gseAdultA_raw))
fit_AdultA <- lmFit(exprs(gseAdultA_raw), design_AdultA)
fit_AdultA <- eBayes(fit_AdultA)

fit_good_rows_A <- suppressMessages(topTable(fit_AdultA, genelist = fDataAdultA_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultA <- gseAdultA_raw[fit_good_rows_A]
```


### Normalization

```{r}
eMatA <- exprs(gseAdultA)
#summary(eMatA)
summary(eMatA[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gseAdultA_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Adult: B. GSE1563

### Basic statistics

```{r}
#pData(gseAdultB_raw)
#colnames(pData(gseAdultB_raw))
gseAdultB_raw$Outcome <- ifelse(grepl("acute rejection", pData(gseAdultB_raw)$description), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultB_raw$Outcome)
dim(gseAdultB_raw)
```

### Cleaning

```{r}
# preprocessing
fDataAdultB_raw <- fData(gseAdultB_raw) # Extract the gene information
colnames(fDataAdultB_raw)
head(fDataAdultB_raw[,"Gene Symbol"])

design_AdultB <- model.matrix(~Outcome, data = pData(gseAdultB_raw))
fit_AdultB <- lmFit(exprs(gseAdultB_raw), design_AdultB)
fit_AdultB <- eBayes(fit_AdultB)

fit_good_rows_B <- suppressMessages(topTable(fit_AdultB, genelist = fDataAdultB_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultB <- gseAdultB_raw[fit_good_rows_B]
```

### Normalization

```{r}
eMatB <- exprs(gseAdultB)
summary(eMatB[,1])
```
This suggests that the data may has been normalised. So we need to normalise the data.

```{r}
# log2 transform - normalisation

#Check if it is normalised(may not be needed)
 qx <- as.numeric(quantile(eMatB, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
 LogC <- (qx[5] > 100) ||
           (qx[6]-qx[1] > 50 && qx[2] > 0)
 if (LogC) { eMatB[which(eMatB <= 0)] <- NaN
   eMatB <- log2(eMatB) }

summary(eMatB[,1])
```

### Boxplot 

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(na.omit(melt(eMatB)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Adult: C. GSE131179

### Basic statistics

```{r}
#pData(gseAdultC_raw)
#colnames(pData(gseAdultC_raw))
gseAdultC_raw$Outcome <- ifelse(grepl("acute rejection", pData(gseAdultC_raw)$characteristics_ch1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultC_raw$Outcome)
dim(gseAdultC_raw)
```

### Cleaning

```{r}
# preprocessing
fData(gseAdultC_raw)
fDataAdultC_raw <- fData(gseAdultC_raw) # Extract the gene information
colnames(fDataAdultC_raw)
head(fDataAdultC_raw[,"Gene Symbol"])

design_AdultC <- model.matrix(~Outcome, data = pData(gseAdultC_raw))
fit_AdultC <- lmFit(exprs(gseAdultC_raw), design_AdultC)
fit_AdultC <- eBayes(fit_AdultC)

fit_good_rows_C <- suppressMessages(topTable(fit_AdultC, genelist = fDataAdultC_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultC <- gseAdultC_raw[fit_good_rows_C]
```

### Normalization

```{r}
eMatC <- exprs(gseAdultC)
summary(eMatC[,1])
```
This suggests that the data may has been normalised. So we need to normalise the data.

```{r}
# log2 transform - normalisation

#Check if it is normalised(may not be needed)
 qx <- as.numeric(quantile(eMatC, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
 LogC <- (qx[5] > 100) ||
           (qx[6]-qx[1] > 50 && qx[2] > 0)
 if (LogC) { eMatC[which(eMatC <= 0)] <- NaN
   eMatC <- log2(eMatC) }

summary(eMatC[,1])
```

### Boxplot 

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(na.omit(melt(eMatC)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

# Merging

```{r}
# changing row names to gene symbol
rownames(eMatA) <- make.unique(fData(gseAdultA)[,"Gene Symbol"])
rownames(eMatB) <- make.unique(fData(gseAdultB)[,"Gene Symbol"])
rownames(eMatC) <- make.unique(fData(gseAdultC)[,"Gene Symbol"])
```


```{r}
# Find common gene
int_adult <- Reduce(intersect, list(rownames(eMatA), rownames(eMatB), rownames(eMatC)))
head(int_adult)
eMatA_common <- eMatA[int_adult,]
eMatB_common <- eMatB[int_adult,]
eMatC_common <- eMatC[int_adult,]

# combine datasets
combined_data <- cbind(eMatA_common, eMatB_common, eMatC_common)

batch <- c(rep("A", ncol(eMatA_common)), 
           rep("B", ncol(eMatB_common)),
           rep("C", ncol(eMatC_common)))

merged_data <- ComBat(dat=combined_data, batch=batch)
```

```{r}
#glimpse at the merged data
head(merged_data)
```

```{r}
as.data.frame(merged_data[,1])

write.csv(as.data.frame(merged_data[,1]), "patient_example.csv", row.names = TRUE)
```


```{r}
#Merge outcome (rejection/stable)
Outcome <- c(gseAdultA_raw$Outcome,gseAdultB_raw$Outcome,gseAdultC_raw$Outcome)
```


# Modeling

## Merged Dataset for Adult

```{r}
X_gse1c = t(merged_data)
y_gse1c = Outcome
```

```{r}
set.seed(3888)

cvK = 5    # Number of CV folds
n_sim = 20 # Number of repeats


# SVM CV for gse1 (biopsy)

cv_accuracy_gse1c = numeric(n_sim) # Vector to store averaged CV accuracies

for (i in 1:n_sim) {
  
  cvSets = cvFolds(nrow(X_gse1c), cvK) # Folds object for cross-validation
  cv_accuracy_folds = numeric(cvK) # Vector to store accuracy for each fold
  
  for (j in 1:cvK) {
    test_id = cvSets$subsets[cvSets$which == j]
    X_train = X_gse1c[-test_id,]
    X_test = X_gse1c[test_id,]
    y_train = y_gse1c[-test_id]
    y_test = y_gse1c[test_id]
    
    design <- model.matrix(~y_train)
    fit <- lmFit(t(X_train), design)
    fit <- eBayes(fit)
    top <- suppressMessages(topTable(fit, n = 50))
    DE_genes <- rownames(top)
    
    X_train = X_train[,DE_genes]
    X_test = X_test[,DE_genes]
    
    svm_fit = svm(x = X_train, y = as.factor(y_train))
    predictions = predict(svm_fit, X_test)
    cv_accuracy_folds[j] = mean(y_test == predictions)
  }
  cv_accuracy_gse1c[i] = mean(cv_accuracy_folds)
}
```



## Plot

```{r}
plot_df = data.frame(accuracy = c(cv_accuracy_gse1c),
                     model = c(rep("Adult", length(cv_accuracy_gse1c))))
                     

ggplot(data = plot_df, aes(x = model, y = accuracy)) +
  geom_boxplot() +
  labs(x = "Model", y = "Mean 5-fold CV accuracy", title = "Distribution of mean 5-fold CV accuracies across\n50 repeats for models") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

# Toptable?

```{r}
featureDataB <- fData(gseAdultB)
#DT::datatable(featureDataA)
designB <- model.matrix(~Outcome, data = pData(gseAdultB))
fitB <- lmFit(eMatB, designB)
fitB <- eBayes(fitB)
topTableB <- suppressMessages(topTable(fitB, genelist=str_split_i(fData(gseAdultB)$"Gene Symbol", " /// ", 1), n = 300))# %>% signif(3))
DT::datatable(topTableB)
```

# Extra comments (Delete later)

SVA Package
merge exprs data

make toptable after merge

check assignment sample solution
