---
title: "External Validation"
author: "KidneyA11"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# load library
#aBiocManager::install("preprocessCore")
library(GEOquery)
library(dplyr)
library(tidyverse)
library(limma)
library(reshape2)
library(DT)
library(stringr)
library(preprocessCore)
library(sva)
library(cvTools)      # Cross-validation
library(e1071)        # Support vector machines, Naive Bayes
library(randomForest) # Random forests
library(viridis)      # Viridis colour scheme
library(plotly)       # Interactive plots
```

# Load Datasets
```{r warning=FALSE}
# load datasets 
# "Adult" for Adult
gseAdultA <- getGEO("GSE46474")
gseAdultA_raw <- gseAdultA[[1]] # Extract the expression data
gseAdultB <- getGEO("GSE1563")
gseAdultB_raw <- gseAdultB[[1]] # Extract the expression data
gseAdultC <- getGEO("GSE14328")
gseAdultC_raw <- gseAdultC[[1]] # Extract the expression data
gseAdultD <- getGEO("GSE21374")
gseAdultD_raw <- gseAdultD[[1]] # Extract the expression data
# "P" for Pediatric
gsePA <- getGEO("GSE20300")
gsePA_raw <- gsePA[[1]] # Extract the expression data
gsePB <- getGEO("GSE174020")
gsePB_raw <- gsePB[[1]] # Extract the expression data
gsePC <- getGEO("GSE75693")
gsePC_raw <- gsePC[[1]] # Extract the expression data
```

# Adult Datasets

## Adult: A. GSE46474

### Basic statistics

```{r}
#pData(gseAdultA_raw)
#colnames(pData(gseAdultA_raw))
gseAdultA_raw$Outcome <- ifelse(grepl("AR", pData(gseAdultA_raw)$characteristics_ch1.5), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultA_raw$Outcome)
dim(gseAdultA_raw)
```

### Cleaning

```{r}
# preprocessing
fDataAdultA_raw <- fData(gseAdultA_raw) # Extract the gene information
#colnames(fDataAdultA_raw)
#head(fDataAdultA_raw[,"Gene Symbol"])
design_AdultA <- model.matrix(~Outcome, data = pData(gseAdultA_raw))
fit_AdultA <- lmFit(exprs(gseAdultA_raw), design_AdultA)
fit_AdultA <- eBayes(fit_AdultA)
fit_good_rows_A <- suppressMessages(topTable(fit_AdultA, genelist = fDataAdultA_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultA <- gseAdultA_raw[fit_good_rows_A]
```


### Normalization

```{r}
eMatA <- exprs(gseAdultA)
#summary(eMatA)
summary(eMatA[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gseAdultA_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Adult: B. GSE1563

### Basic statistics

```{r}
#pData(gseAdultB_raw)
#colnames(pData(gseAdultB_raw))
gseAdultB_raw$Outcome <- ifelse(grepl("acute rejection", pData(gseAdultB_raw)$description), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultB_raw$Outcome)
dim(gseAdultB_raw)
```

### Cleaning

```{r}
# preprocessing
fDataAdultB_raw <- fData(gseAdultB_raw) # Extract the gene information
colnames(fDataAdultB_raw)
head(fDataAdultB_raw[,"Gene Symbol"])
design_AdultB <- model.matrix(~Outcome, data = pData(gseAdultB_raw))
fit_AdultB <- lmFit(exprs(gseAdultB_raw), design_AdultB)
fit_AdultB <- eBayes(fit_AdultB)
fit_good_rows_B <- suppressMessages(topTable(fit_AdultB, genelist = fDataAdultB_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultB <- gseAdultB_raw[fit_good_rows_B]
```

### Normalization

```{r}
eMatB <- exprs(gseAdultB)
summary(eMatB[,1])
```
This suggests that the data may has been normalised. So we need to normalise the data.

```{r}
# log2 transform - normalisation
#Check if it is normalised(may not be needed)
 qx <- as.numeric(quantile(eMatB, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
 LogC <- (qx[5] > 100) ||
           (qx[6]-qx[1] > 50 && qx[2] > 0)
 if (LogC) { eMatB[which(eMatB <= 0)] <- NaN
   eMatB <- log2(eMatB) }
summary(eMatB[,1])
```

### Boxplot 

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(na.omit(melt(eMatB)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Adult: C. GSE131179

### Basic statistics

```{r}
pData(gseAdultC_raw)
colnames(pData(gseAdultC_raw))
gseAdultC_raw$Outcome <- ifelse(grepl("acute rejection", pData(gseAdultC_raw)$characteristics_ch1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gseAdultC_raw$Outcome)
dim(gseAdultC_raw)
```

### Cleaning

```{r}
# preprocessing
fData(gseAdultC_raw)
fDataAdultC_raw <- fData(gseAdultC_raw) # Extract the gene information
colnames(fDataAdultC_raw)
head(fDataAdultC_raw[,"Gene Symbol"])
design_AdultC <- model.matrix(~Outcome, data = pData(gseAdultC_raw))
fit_AdultC <- lmFit(exprs(gseAdultC_raw), design_AdultC)
fit_AdultC <- eBayes(fit_AdultC)
fit_good_rows_C <- suppressMessages(topTable(fit_AdultC, genelist = fDataAdultC_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultC <- gseAdultC_raw[fit_good_rows_C]
```

### Normalization

```{r}
eMatC <- exprs(gseAdultC)
summary(eMatC[,1])
```
This suggests that the data may has been normalised. So we need to normalise the data.

```{r}
# log2 transform - normalisation
#Check if it is normalised(may not be needed)
 qx <- as.numeric(quantile(eMatC, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
 LogC <- (qx[5] > 100) ||
           (qx[6]-qx[1] > 50 && qx[2] > 0)
 if (LogC) { eMatC[which(eMatC <= 0)] <- NaN
   eMatC <- log2(eMatC) }
summary(eMatC[,1])
```

## Adult: D. GSE21374

### Basic statistics

```{r}
pData(gseAdultD_raw)
colnames(pData(gseAdultD_raw))
gseAdultD_raw$Outcome <- ifelse(grepl("nonrej", pData(gseAdultD_raw)$characteristics_ch1.3), "Stable", "Rejection")

 #Tidy the title variable and call it Outcome
table(gseAdultD_raw$Outcome)
dim(gseAdultD_raw)
```

### Cleaning

```{r}
# preprocessing
fData(gseAdultD_raw)
fDataAdultD_raw <- fData(gseAdultD_raw) # Extract the gene information
colnames(fDataAdultD_raw)
head(fDataAdultD_raw[,"Gene Symbol"])
design_AdultD <- model.matrix(~Outcome, data = pData(gseAdultD_raw))
fit_AdultD <- lmFit(exprs(gseAdultD_raw), design_AdultD)
fit_AdultD <- eBayes(fit_AdultD)
fit_good_rows_D <- suppressMessages(topTable(fit_AdultD, genelist = fDataAdultD_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gseAdultD <- gseAdultD_raw[fit_good_rows_D]
```

### Normalization

```{r}
eMatD <- exprs(gseAdultD)
summary(eMatD[,1])
```
This suggests that the data may has been normalised. So we need to normalise the data.

```{r}
# log2 transform - normalisation
#Check if it is normalised(may not be needed)
 qx <- as.numeric(quantile(eMatD, c(0., 0.25, 0.5, 0.75, 0.99, 1.0), na.rm=T))
 LogC <- (qx[5] > 100) ||
           (qx[6]-qx[1] > 50 && qx[2] > 0)
 if (LogC) { eMatD[which(eMatD <= 0)] <- NaN
   eMatD <- log2(eMatD) }
summary(eMatD[,1])
```

### Boxplot 

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(na.omit(melt(eMatD)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```
# Pediatric Datasets

## Pediatric: A. GSE20300

### Basic statistics

```{r}
#pData(gsePA_raw)
#colnames(pData(gsePA_raw))
gsePA_raw$Outcome <- ifelse(grepl("Stable", pData(gsePA_raw)$characteristics_ch1), "Stable", "Rejection") #Tidy the title variable and call it Outcome
table(gsePA_raw$Outcome)
dim(gsePA_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPA_raw <- fData(gsePA_raw) # Extract the gene information
#colnames(fDataPA_raw)
#head(fDataPA_raw[,"Gene Symbol"])
design_PA <- model.matrix(~Outcome, data = pData(gsePA_raw))
fit_PA <- lmFit(exprs(gsePA_raw), design_PA)
fit_PA <- eBayes(fit_PA)
fit_good_rows_PA <- suppressMessages(topTable(fit_PA, genelist = fDataPA_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePA <- gsePA_raw[fit_good_rows_PA]
```


### Normalization

```{r}
eMatPA <- exprs(gsePA)
#summary(eMatPA)
summary(eMatPA[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePA_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Pediatric: B. GSE174020

### Basic statistics

```{r}
#pData(gsePB_raw)
#colnames(pData(gsePB_raw))
gsePB_raw$Outcome <- ifelse(grepl("AR", pData(gsePB_raw)$characteristics_ch1.1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gsePB_raw$Outcome)
dim(gsePB_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPB_raw <- fData(gsePB_raw) # Extract the gene information
#colnames(fDataPB_raw)
head(fDataPB_raw[,"Gene Symbol"])
design_PB <- model.matrix(~Outcome, data = pData(gsePB_raw))
fit_PB <- lmFit(exprs(gsePB_raw), design_PB)
fit_PB <- eBayes(fit_PB)
fit_good_rows_PB <- suppressMessages(topTable(fit_PB, genelist = fDataPB_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePB <- gsePB_raw[fit_good_rows_PB]
```


### Normalization

```{r}
eMatPB <- exprs(gsePB)
#summary(eMatPB)
summary(eMatPB[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePB_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Pediatric: C. GSE75693

### Basic statistics

```{r}
pData(gsePC_raw)
#colnames(pData(gsePC_raw))
gsePC_raw$Outcome <- ifelse(grepl("Acute rejection", pData(gsePC_raw)$characteristics_ch1.1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gsePC_raw$Outcome)
dim(gsePC_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPC_raw <- fData(gsePC_raw) # Extract the gene information
#colnames(fDataPC_raw)
#head(fDataPC_raw[,"Gene Symbol"])
design_PC <- model.matrix(~Outcome, data = pData(gsePC_raw))
fit_PC <- lmFit(exprs(gsePC_raw), design_PC)
fit_PC <- eBayes(fit_PC)
fit_good_rows_PC <- suppressMessages(topTable(fit_PC, genelist = fDataPC_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePC <- gsePC_raw[fit_good_rows_PC]
```


### Normalization

```{r}
eMatPC <- exprs(gsePC)
#summary(eMatPC)
summary(eMatPC[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePC_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```


# Merging

## Adult

```{r}
# changing row names to gene symbol
rownames(eMatA) <- make.unique(fData(gseAdultA)[,"Gene Symbol"])
rownames(eMatB) <- make.unique(fData(gseAdultB)[,"Gene Symbol"])
rownames(eMatC) <- make.unique(fData(gseAdultC)[,"Gene Symbol"])
rownames(eMatD) <- make.unique(fData(gseAdultD)[,"Gene Symbol"])
```


```{r}
# Find common gene
int_adult <- Reduce(intersect, list(rownames(eMatA), rownames(eMatB), rownames(eMatC)))
head(int_adult)
eMatA_common <- eMatA[int_adult,]
eMatB_common <- eMatB[int_adult,]
eMatC_common <- eMatC[int_adult,]
# combine datasets
combined_data <- cbind(eMatA_common, eMatB_common, eMatC_common)
batch <- c(rep("A", ncol(eMatA_common)), 
           rep("B", ncol(eMatB_common)),
           rep("C", ncol(eMatC_common)))
merged_dataA <- ComBat(dat=combined_data, batch=batch)
```

```{r}
#glimpse at the merged data
head(merged_dataA)
```

```{r}
as.data.frame(merged_dataA[,1])
write.csv(as.data.frame(merged_dataA[,1]), "patient_example.csv", row.names = TRUE)
```



```{r}
#Merge outcome (rejection/stable)
OutcomeA <- c(gseAdultA_raw$Outcome,gseAdultB_raw$Outcome,gseAdultC_raw$Outcome)
```


## Pediatric
```{r}
# changing row names to gene symbol
rownames(eMatPA) <- make.unique(fData(gsePA)[,"Gene Symbol"])
rownames(eMatPB) <- make.unique(fData(gsePB)[,"Gene Symbol"])
rownames(eMatPC) <- make.unique(fData(gsePC)[,"Gene Symbol"])
```


```{r}
# Find common gene
int_P <- Reduce(intersect, list(rownames(eMatPA), rownames(eMatPB), rownames(eMatPC)))
head(int_P)
eMatPA_common <- eMatPA[int_P,]
eMatPB_common <- eMatPB[int_P,]
eMatPC_common <- eMatPC[int_P,]
# combine datasets
combined_data <- cbind(eMatPA_common, eMatPB_common, eMatPC_common)
batch <- c(rep("A", ncol(eMatPA_common)), 
           rep("B", ncol(eMatPB_common)),
           rep("C", ncol(eMatPC_common)))
merged_dataP <- ComBat(dat=combined_data, batch=batch)
```

```{r}
#glimpse at the merged data
head(merged_dataP)
```


```{r}
#Merge outcome (rejection/stable)
OutcomeP <- c(gsePA_raw$Outcome,gsePB_raw$Outcome,gsePC_raw$Outcome)
```



# Modeling

## Merged Dataset for Adult

```{r}
X_gse1c = t(eMatD)
y_gse1c = gseAdultD_raw$Outcome
```

```{r}
set.seed(3888)
cvK = 5    # Number of CV folds
n_sim = 20 # Number of repeats
# SVM CV for gse1 (biopsy)
cv_accuracy_gse1c = numeric(n_sim) # Vector to store averaged CV accuracies for SVM
cv_accuracy_gse1c_rf = numeric(n_sim) # Vector to store averaged CV accuracies for Random Forest
cv_accuracy_gse1c_nb = numeric(n_sim) # Vector to store averaged CV accuracies for Naive Bayes
for (i in 1:n_sim) {
  
  cvSets = cvFolds(nrow(X_gse1c), cvK) # Folds object for cross-validation
  cv_accuracy_folds = numeric(cvK) # Vector to store accuracy for each fold for SVM
  cv_accuracy_folds_rf = numeric(cvK) # Vector to store accuracy for each fold for Random Forest
  cv_accuracy_folds_nb = numeric(cvK) # Vector to store accuracy for each fold for Naive Bayes
  
  
  for (j in 1:cvK) {
    test_id = cvSets$subsets[cvSets$which == j]
    X_train = X_gse1c[-test_id,]
    X_test = X_gse1c[test_id,]
    y_train = y_gse1c[-test_id]
    y_test = y_gse1c[test_id]
    
    design <- model.matrix(~y_train)
    fit <- lmFit(t(X_train), design)
    fit <- eBayes(fit)
    top <- suppressMessages(topTable(fit, n = 50))
    DE_genes <- rownames(top)
    
    X_train = X_train[,DE_genes]
    X_test = X_test[,DE_genes]
    
    # SVM model
    svm_fit = svm(x = X_train, y = as.factor(y_train))
    predictions = predict(svm_fit, X_test)
    cv_accuracy_folds[j] = mean(y_test == predictions)
    
    # Random Forest model
    rf_fit = randomForest(x = X_train, y = as.factor(y_train))
    predictions_rf = predict(rf_fit, X_test)
    cv_accuracy_folds_rf[j] = mean(y_test == predictions_rf)
    
    # Naive Bayes model
    nb_fit = naiveBayes(x = X_train, y = as.factor(y_train))
    predictions_nb = predict(nb_fit, X_test)
    cv_accuracy_folds_nb[j] = mean(y_test == predictions_nb)
  }
  cv_accuracy_gse1c[i] = mean(cv_accuracy_folds)
  cv_accuracy_gse1c_rf[i] = mean(cv_accuracy_folds_rf)
  cv_accuracy_gse1c_nb[i] = mean(cv_accuracy_folds_nb)
}
```

### Predict

```{r}
X_gse1c <- t(merged_dataA)
y_gse1c <- OutcomeA
svm_fit <- svm(x = X_gse1c, y = as.factor(y_gse1c), probability = TRUE)
patient_data <- t(merged_dataA[,5]) 
prediction <- predict(svm_fit, newdata = as.data.frame(patient_data), probability = TRUE)
prediction
OutcomeA[5]
```

## Merged Dataset for Pediatric

```{r}
X_gse2c = t(merged_dataP)
y_gse2c = OutcomeP
```

```{r}
set.seed(3888)
cvK = 5    # Number of CV folds
n_sim = 20 # Number of repeats
# SVM CV for gse2 (biopsy)
cv_accuracy_gse2c = numeric(n_sim) # Vector to store averaged CV accuracies for SVM
cv_accuracy_gse2c_rf = numeric(n_sim) # Vector to store averaged CV accuracies for Random Forest
cv_accuracy_gse2c_nb = numeric(n_sim) # Vector to store averaged CV accuracies for Naive Bayes
for (i in 1:n_sim) {
  
  cvSets = cvFolds(nrow(X_gse2c), cvK) # Folds object for cross-validation
  cv_accuracy_folds2 = numeric(cvK) # Vector to store accuracy for each fold V
  cv_accuracy_folds_rf2 = numeric(cvK) # Vector to store accuracy for each fold for Random Forest
  cv_accuracy_folds_nb2 = numeric(cvK) # Vector to store accuracy for each fold for Naive Bayes
  
  
  for (j in 1:cvK) {
    test_id = cvSets$subsets[cvSets$which == j]
    X_trainP = X_gse2c[-test_id,]
    X_testP = X_gse2c[test_id,]
    y_trainP = y_gse2c[-test_id]
    y_testP = y_gse2c[test_id]
    
    designP <- model.matrix(~y_trainP)
    fitP <- lmFit(t(X_trainP), designP)
    fitP <- eBayes(fitP)
    top_P <- suppressMessages(topTable(fitP, n = 50))
    DE_genes_P <- rownames(top_P)
    
    X_trainP = X_trainP[,DE_genes_P]
    X_testP = X_testP[,DE_genes_P]
    
    # SVM model
    svm2_fit = svm(x = X_trainP, y = as.factor(y_trainP))
    predictions2 = predict(svm2_fit, X_testP)
    cv_accuracy_folds2[j] = mean(y_testP == predictions2)
    
    # Random Forest model
    rf2_fit = randomForest(x = X_trainP, y = as.factor(y_trainP))
    predictions_rf2 = predict(rf2_fit, X_testP)
    cv_accuracy_folds_rf2[j] = mean(y_testP == predictions_rf2)
    
    # Naive Bayes model
    nb2_fit = naiveBayes(x = X_trainP, y = as.factor(y_trainP))
    predictions_nb2 = predict(nb2_fit, X_testP)
    cv_accuracy_folds_nb2[j] = mean(y_testP == predictions_nb2)
  }
  cv_accuracy_gse2c[i] = mean(cv_accuracy_folds2)
  cv_accuracy_gse2c_rf[i] = mean(cv_accuracy_folds_rf2)
  cv_accuracy_gse2c_nb[i] = mean(cv_accuracy_folds_nb2)
}
```

### Predict

```{r}
X_gse2c <- t(merged_dataP)
y_gse2c <- OutcomeP
svm_fit_P <- svm(x = X_gse2c, y = as.factor(y_gse2c), probability = TRUE)
patient_dataP <- t(merged_dataP[,5]) 
predictionP <- predict(svm_fit_P, newdata = as.data.frame(patient_dataP), probability = TRUE)
predictionP
OutcomeP[5]
```


## Plot

```{r}
plot_df = data.frame(accuracy = c(cv_accuracy_gse1c,
                                  cv_accuracy_gse1c_rf,
                                  cv_accuracy_gse1c_nb,
                                  cv_accuracy_gse2c,
                                  cv_accuracy_gse2c_rf,
                                  cv_accuracy_gse2c_nb),
                     model = c(rep("Adult-SVM", length(cv_accuracy_gse1c)),
                               rep("Adult-RF", length(cv_accuracy_gse1c_rf)),
                               rep("Adult-NB", length(cv_accuracy_gse1c_nb)),
                               rep("Pediatric-SVM", length(cv_accuracy_gse2c)),
                               rep("Pediatric-RF", length(cv_accuracy_gse2c_rf)),
                               rep("Pediatric-NB", length(cv_accuracy_gse2c_nb))))
                     
ggplot(data = plot_df, aes(x = model, y = accuracy)) +
  geom_boxplot() +
  labs(x = "Model", y = "Mean 5-fold CV accuracy", title = "Distribution of mean 5-fold CV accuracies across\n50 repeats for models") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

# Toptable?

```{r}
featureDataB <- fData(gseAdultB)
#DT::datatable(featureDataA)
designB <- model.matrix(~Outcome, data = pData(gseAdultB))
fitB <- lmFit(eMatB, designB)
fitB <- eBayes(fitB)
topTableB <- suppressMessages(topTable(fitB, genelist=str_split_i(fData(gseAdultB)$"Gene Symbol", " /// ", 1), n = 300))# %>% signif(3))
DT::datatable(topTableB)
```

# Extra comments (Delete later)

SVA Package
merge exprs data

make toptable after merge

check assignment sample solution
