---
title: "Paediatric Dataset"
author: "KidneyA11"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE}
# load library
#aBiocManager::install("preprocessCore")
library(GEOquery)
library(dplyr)
library(tidyverse)
library(limma)
library(reshape2)
library(DT)
library(stringr)
library(preprocessCore)
library(sva)
library(cvTools)      # Cross-validation
library(e1071)        # Support vector machines
library(randomForest) # Random forests
library(viridis)      # Viridis colour scheme
library(plotly)       # Interactive plots
```

# Load Datasets
```{r warning=FALSE}
# "P" for Paediatric

# load datasets for Paediatric

gsePA <- getGEO("GSE20300")
gsePA_raw <- gsePA[[1]] # Extract the expression data

gsePB <- getGEO("GSE174020")
gsePB_raw <- gsePB[[1]] # Extract the expression data

gsePC <- getGEO("GSE75693")
gsePC_raw <- gsePC[[1]] # Extract the expression data
```

## Paediatric: A. GSE20300

### Basic statistics

```{r}
#pData(gsePA_raw)
#colnames(pData(gsePA_raw))
gsePA_raw$Outcome <- ifelse(grepl("Stable", pData(gsePA_raw)$characteristics_ch1), "Stable", "Rejection") #Tidy the title variable and call it Outcome
table(gsePA_raw$Outcome)
dim(gsePA_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPA_raw <- fData(gsePA_raw) # Extract the gene information
#colnames(fDataPA_raw)
#head(fDataPA_raw[,"Gene Symbol"])

design_PA <- model.matrix(~Outcome, data = pData(gsePA_raw))
fit_PA <- lmFit(exprs(gsePA_raw), design_PA)
fit_PA <- eBayes(fit_PA)

fit_good_rows_PA <- suppressMessages(topTable(fit_PA, genelist = fDataPA_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePA <- gsePA_raw[fit_good_rows_PA]
```


### Normalization

```{r}
eMatPA <- exprs(gsePA)
#summary(eMatPA)
summary(eMatPA[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePA_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Paediatric: B. GSE174020

### Basic statistics

```{r}
#pData(gsePB_raw)
#colnames(pData(gsePB_raw))
gsePB_raw$Outcome <- ifelse(grepl("AR", pData(gsePB_raw)$characteristics_ch1.1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gsePB_raw$Outcome)
dim(gsePB_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPB_raw <- fData(gsePB_raw) # Extract the gene information
#colnames(fDataPB_raw)
#head(fDataPB_raw[,"Gene Symbol"])

design_PB <- model.matrix(~Outcome, data = pData(gsePB_raw))
fit_PB <- lmFit(exprs(gsePB_raw), design_PB)
fit_PB <- eBayes(fit_PB)

fit_good_rows_PB <- suppressMessages(topTable(fit_PB, genelist = fDataPB_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePB <- gsePB_raw[fit_good_rows_PB]
```


### Normalization

```{r}
eMatPB <- exprs(gsePB)
#summary(eMatPB)
summary(eMatPB[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePB_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

## Paediatric: C. GSE75693

### Basic statistics

```{r}
pData(gsePC_raw)
#colnames(pData(gsePC_raw))
gsePC_raw$Outcome <- ifelse(grepl("Acute rejection", pData(gsePC_raw)$characteristics_ch1.1), "Rejection", "Stable") #Tidy the title variable and call it Outcome
table(gsePC_raw$Outcome)
dim(gsePC_raw)
```

### Cleaning

```{r warning=FALSE}
# preprocessing
fDataPC_raw <- fData(gsePC_raw) # Extract the gene information
#colnames(fDataPC_raw)
#head(fDataPC_raw[,"Gene Symbol"])

design_PC <- model.matrix(~Outcome, data = pData(gsePC_raw))
fit_PC <- lmFit(exprs(gsePC_raw), design_PC)
fit_PC <- eBayes(fit_PC)

fit_good_rows_PC <- suppressMessages(topTable(fit_PC, genelist = fDataPC_raw[, "Gene Symbol"], n = Inf)) %>%
  rownames_to_column("row") %>%
  filter(!is.na(ID)) %>%
  filter(ID != "") %>%
  group_by(ID) %>%
  filter(P.Value == min(P.Value)) %>%
  pull(row)
gsePC <- gsePC_raw[fit_good_rows_PC]
```


### Normalization

```{r}
eMatPC <- exprs(gsePC)
#summary(eMatPC)
summary(eMatPC[,1])
```
This suggests that the data may has been normalised. 

### Boxplot

```{r}
## base R code is
## boxplot(eMat) or boxplot(melt(exprs(gse)))
p <- ggplot(melt(exprs(gsePC_raw)), aes(x=Var2, y=value)) +  
  geom_boxplot(outlier.colour="black", outlier.shape=16,outlier.size=0.5, notch=FALSE) +
  theme(axis.text.x = element_text(angle = 180, hjust = 1)) +
  labs (x = "patient", y = "expression value") + theme_minimal()
p
```

# Merging

```{r}
# changing row names to gene symbol
rownames(eMatPA) <- make.unique(fData(gsePA)[,"Gene Symbol"])
rownames(eMatPB) <- make.unique(fData(gsePB)[,"Gene Symbol"])
rownames(eMatPC) <- make.unique(fData(gsePC)[,"Gene Symbol"])
```


```{r}
# Find common gene
int_P <- Reduce(intersect, list(rownames(eMatPA), rownames(eMatPB), rownames(eMatPC)))
head(int_P)
eMatPA_common <- eMatPA[int_P,]
eMatPB_common <- eMatPB[int_P,]
eMatPC_common <- eMatPC[int_P,]

# combine datasets
combined_data <- cbind(eMatPA_common, eMatPB_common, eMatPC_common)

batch <- c(rep("A", ncol(eMatPA_common)), 
           rep("B", ncol(eMatPB_common)),
           rep("C", ncol(eMatPC_common)))

merged_data <- ComBat(dat=combined_data, batch=batch)
```

```{r}
#glimpse at the merged data
head(merged_data)
```


```{r}
#Merge outcome (rejection/stable)
Outcome <- c(gsePA_raw$Outcome,gsePB_raw$Outcome,gsePC_raw$Outcome)
```


# Modeling

## Merged Dataset for Paediatric

```{r}
X_gse2c = t(merged_data)
y_gse2c = Outcome
```

```{r}
set.seed(3888)

cvK = 5    # Number of CV folds
n_sim = 20 # Number of repeats


# SVM CV for gse1 (biopsy)

cv_accuracy_gse2c = numeric(n_sim) # Vector to store averaged CV accuracies

for (i in 1:n_sim) {
  
  cvSets = cvFolds(nrow(X_gse2c), cvK) # Folds object for cross-validation
  cv_accuracy_folds = numeric(cvK) # Vector to store accuracy for each fold
  
  for (j in 1:cvK) {
    test_id = cvSets$subsets[cvSets$which == j]
    X_train = X_gse2c[-test_id,]
    X_test = X_gse2c[test_id,]
    y_train = y_gse2c[-test_id]
    y_test = y_gse2c[test_id]
    
    design <- model.matrix(~y_train)
    fit <- lmFit(t(X_train), design)
    fit <- eBayes(fit)
    top <- suppressMessages(topTable(fit, n = 50))
    DE_genes <- rownames(top)
    
    X_train = X_train[,DE_genes]
    X_test = X_test[,DE_genes]
    
    svm_fit = svm(x = X_train, y = as.factor(y_train))
    predictions = predict(svm_fit, X_test)
    cv_accuracy_folds[j] = mean(y_test == predictions)
  }
  cv_accuracy_gse2c[i] = mean(cv_accuracy_folds)
}
```



## Plot

```{r}
plot_df = data.frame(accuracy = c(cv_accuracy_gse2c),
                     model = c(rep("Paediatric", length(cv_accuracy_gse2c))))
                     

ggplot(data = plot_df, aes(x = model, y = accuracy)) +
  geom_boxplot() +
  labs(x = "Model", y = "Mean 5-fold CV accuracy", title = "Distribution of mean 5-fold CV accuracies across\n50 repeats for models") +
  theme(plot.title = element_text(hjust = 0.5, size = 12),
        panel.border = element_rect(colour = "black", fill = NA, size = 1))
```

# Extra comments (Delete later)

SVA Package
merge exprs data

make toptable after merge

check assignment sample solution
